### 2020-02-28 ChapterBackground/chapterBackground.tex
Added: Consistency in Real-Time Physics

\subsection{Dead-Reckoning}

@@

\section{Physics in Online Gaming}
%(and the lack of, due to the challenges faced). Mostly limited to dead-reckoning
%https://books.google.co.uk/books?hl=en&lr=&id=ujfOBQAAQBAJ&oi=fnd&pg=PA307&dq=Physics+networked+games&ots=OC5YWxmd25&sig=vcxatd9Wl0_SdxCAUZvmP6BWl7I&redir_esc=y#v=onepage&q=Physics%20networked%20games&f=false

@@

\section{Graph-Partitioning}
%When it comes to spatially-partitioning a simulation, the problem is essentially a graph partitioning problem, where objects are nodes and if there’s a possible interaction, an edge exists. Useful for terms, comparisons with and future work on load-balancing will heavily involve this.

>>>>

@@

\section{Consistency in Real-Time Physics}
%(and the lack of, due to the challenges faced). Mostly limited to dead-reckoning
%https://books.google.co.uk/books?hl=en&lr=&id=ujfOBQAAQBAJ&oi=fnd&pg=PA307&dq=Physics+networked+games&ots=OC5YWxmd25&sig=vcxatd9Wl0_SdxCAUZvmP6BWl7I&redir_esc=y#v=onepage&q=Physics%20networked%20games&f=false

@@

Maintaining consistency in the physics aspect of games when distributed is great challenge, this is due to the time-sensitive nature of physics (millisecond differences can result in completely divergent results, discussed further in \ref{divergingResults}) \cite{2012EnhancedDeadReckoning}.

Consistency in physics can be divided into two categories:
\begin{itemize}
\item Consistency in simulation view, for example, differences between the physics simulation on the server and a client.
\item Consistency within the simulation itself (when the simulation is distributed between devices).
\end{itemize}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figs/DeadReckoning}
	\caption{An example of dead-reckoning \cite{2012EnhancedDeadReckoning}.}
	\label{fig_DeadReckoning}
\end{figure}

The former is a very common issue in online gaming, particularly due to high latency and limited bandwidth between servers and client, necessitating the need for client-side correction techniques. The most common technique used is dead-reckoning, shown in Fig. \ref{fig_DeadReckoning}. Dead-reckoning's use in games is well researched, including the suitability for different genres \cite{2002DeadReckoningSuitability}, the accuracy of dead-reckoning \cite{2004AccuracyDeadReckoning}, its effects on fairness \cite{2005FairnessDeadReckoning} and enhancements of the technique, such as \cite{2012EnhancedDeadReckoning} which takes into account environmental objects when dead-reckoning is performed.

Although these correction techniques may be suitable for handling inconsistencies in simulation view between servers and clients, they are not suitable for maintaining consistency within a distributed simulation, which is the focus of this study. This is because in order to prevent diverging results, strong consistency of collisions must be maintained (discussed further in \ref{divergingResults}).

Gameplay aspects of games, such as life/death and score decisions require a high-level of consistency in order to maintain fairness \cite{savery2014consistency}. As a result of the inconsistencies in distributed real-time physics and importance of consistency in gameplay, this leaves game developers with two choices:
\begin{enumerate}
	\item Don't distribute the simulation of the physics, meaning the server(s) simulate the physics for the game (removing inconsistencies within the simulation, but not inconsistencies between the server and client). However, this uses up limited computational resources in single server systems or in the case of multi-server systems, creating complex problems at the boundaries between servers. Problems include not only the handing off of physical entities between servers, but also the interaction of entities between servers and inconsistencies between simulating servers.
	\item Allow physics to be simulated on clients, distributing physics between clients, creating consistency problems. Due to the physics consistency problems, physics would be unable to influence gameplay, otherwise sacrifices to fairness would need be made.
\end{enumerate}

This study aims to address the problems of consistency in distributing real-time physics between servers in multi-server networked games. This would allow for physics to play a larger role in game, including have large influences on gameplay in networked games.
